# Project Progress

## Get and extract data

```{r}
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
downloadDir <- "data"

path <- function(...) { paste(..., sep = "/") }

zipFile <- path(downloadDir, "dataset.zip")
if(!file.exists(zipFile)) { download.file(url, zipFile, method = "curl") }

dataDir <- path(downloadDir, "UCI HAR Dataset")
if(!file.exists(dataDir)) { unzip(zipFile, exdir = downloadDir) }
```


## "1. Merges the training and the test sets to create one data set."

Merge and add in the feature names:

```{r}
read <- function(path) { read.table(path(dataDir, path)) }

if(!exists("XTrain")) { XTrain <- read("train/X_train.txt") }
if(!exists("XTest"))  { XTest  <- read("test/X_test.txt") }
merged <- rbind(XTrain, XTest)

featureNames <- read("features.txt")[, 2]
names(merged) <- featureNames
```

## "2. Extracts only the measurements on the mean and standard deviation for each measurement."

Limit to columns with feature names matching mean() or std():

```{r}
matches <- grep("(mean|std)\\(\\)", names(merged))
limited <- merged[, matches]
```

## "3. Uses descriptive activity names to name the activities in the data set"

Get the activity data and map to nicer names:

```{r}
yTrain <- read("train/y_train.txt")
yTest  <- read("test/y_test.txt")
yMerged <- rbind(yTrain, yTest)[, 1]

activityNames <-
  c("Walking", "Walking Upstairs", "Walking Downstairs", "Sitting", "Standing", "Laying")
activities <- activityNames[yMerged]
```

## "4. Appropriately labels the data set with descriptive variable names."

Change t to Time-, f to Frequency-, mean() to Mean and std() to Standard-Deviation

```{r}
names(limited) <- gsub("^t", "Time", names(limited))
names(limited) <- gsub("^f", "Frequency", names(limited))
names(limited) <- gsub("-mean\\(\\)", "Mean", names(limited))
names(limited) <- gsub("-std\\(\\)", "StdDev", names(limited))
names(limited) <- gsub("-", "", names(limited))
# Some of the data is misnamed as fBodyBody..., fix it up to have only one Body
names(limited) <- gsub("BodyBody", "Body", names(limited))
```


## Add activities and subject with nice names

```{r}
subjectTrain <- read("train/subject_train.txt")
subjectTest  <- read("test/subject_test.txt")
subjects <- rbind(subjectTrain, subjectTest)[, 1]

tidy <- cbind(Subject = subjects, Activity = activities, limited)
```

## "5. Creates a second, independent tidy data set with the average of each variable for each activity and each subject."

```{r}
library(plyr)
# Column means for all but the subject and activity columns
limitedColMeans <- function(data) { colMeans(data[,-c(1,2)]) }
tidyMeans <- ddply(tidy, .(Subject, Activity), limitedColMeans)
names(tidyMeans)[-c(1,2)] <- paste0("Mean", names(tidyMeans)[-c(1,2)])
```

## Write file

```{r}
write.table(tidyMeans, "tidyMeans.txt", row.names = FALSE)
tidyMeansCheck <- read.table("tidyMeans.txt", header = TRUE)
```


